Generating Datasets about Building Designs \\Based on a Multilayer Building Ontology

### 2015-04-30 article-01-introduction.tex
Updates from ShareLaTeX
Edited by Seppo

\begin{figure}[h]
\centering
\includegraphics[angle=0,width=0.9\textwidth]{images/ifcOWL-multilayers.png}
%\includegraphics{images/ifcOWL-multilayers.jpg}
\caption{Multiple layers of \ifcowl{} ontology}
\label{fig:ifcOWL-layers}
\end{figure}

@@

From the previous IFC-to-OWL/RDF conversions
\cite{beetz2005ontology,beetz2009ifcowl,pauwels2011interoperability} our approach differs since it
is based on OWL 2 profiles, produces a flexible multilayer ontology, applies \emph{very different methods of data\-type definitions} and is specifically targeted to
real-life Linked Data applications. It uses stable URIs for entities, in contrast to the previous approaches
where the URIs were instable across models versions.

>>>>

\begin{figure}[h]
\centering
\includegraphics[angle=0,width=1.0\textwidth]{images/ifcOWL-multilayers.png}
%\includegraphics{images/ifcOWL-multilayers.jpg}
\caption{Multiple layers of \ifcowl{} ontology}
\label{fig:ifcOWL-layers}
\end{figure}

@@

From the previous IFC-to-OWL/RDF conversions
\cite{beetz2005ontology,beetz2009ifcowl,pauwels2011interoperability} our approach differs since it
is based on OWL 2 profiles, produces a flexible multilayer ontology, applies other methods of data\-type definitions and is specifically targeted to
real-life Linked Data applications. It uses stable URIs for entities, in contrast to the previous approaches
where the URIs were instable across models versions.

### 2015-04-30 article-02-main.tex
Edited by Seppo

\section{Generating ifcOWL Ontologies And ifcRDF Datasets}
\label{sec:ifcOWL}
This section describes the derivation of the ifcOWL ontologies from the IFC schema. The schema defines series of datatypes, functions and rules by using EXPRESS data definition language (ISO10303-11 \cite{schenck1994information}).

\subsection{Common Principles}

Layer \ifcsimple{} has been designed to be compatible with with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other Semantic Web tools. The other layers \standard{} and \extended{} are compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports representational constructs of EXPRESS, such as enumerations of individuals, union classes and anonymous individuals \cite{motik2012owl}. All \ifcowl{} layers include all type definitions and hierarchies so that any \ifcrdf{} datasets converted from IFC data files could reuse types without defining new ones.

The function and rule definitions, type constraints and derived attributes specified in IFC schema after keywords \name{FUNCTION}, \name{RULE}, and \name{WHERE} are not translated into any \ifcowl{} ontology either because of their procedural nature, or the complexity of the resulting constraints.

@@

All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g. \name{IFC4\_ADD1}):

@@

Types \name{STRING} and \name{BINARY} data\-types may be specified with a size parameter: \name{STRING(max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}. In layers \simple{} and \standard{} the size should be ignored, while in layer \extended{} it should be declared:

@@

The enumeration values cannot be defined as literals because enumerations of literals is not supported by OWL 2 RL. Two different approaches are used because enumerations with more than one individual or literal are not supported in OWL 2 EL or QL. The fact that some values -- for instance, \name{ifc:NOT\-DE\-FI\-NED} -- belong to different enumeration data\-types should not cause any problems since any enumeration value is always specified with an exact type.

@@

\subsubsection{Select Datatypes} represent unions of other data\-types. Particular members of a select data\-type can be entity, defined or select data\-types. For instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}. A select data\-type is converted to a subclass of \name{expr:\-Select\-Class}. Its particular data\-types are defined as subclasses of the select class, but in different ways according to the layer:

@@

Defined data\-types are declared as \name{owl:Class}es using boxing technique because they should be able to merge with other \name{owl:Class}es for creating select data\-types. In addition, values of defined data\-types are qualified with type name. For example, value \name{"5.0"^^xsd:double} for \name{IfcLengthMeasure} is different to \name{IfcMassMeasure} although they both are \name{REAL} and belong to select class \name{IfcMeasureValue}.

In principle type \name{Ifc\-Time\-Stamp} wraps an integer value which represents so-called Unix time. However, we have chosen the type \name{xsd:date\-Time} for these values, and all values of \name{Ifc\-Time\-Stamp} are converted to a standard datetime values.

@@

\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages. Each type may have one super\-type. An attribute is defined with a data\-type, cardinalities and can be marked as optional. An entity data\-type should be converted to an \name{owl:\-Class} which is subclass of its superclass (if any) or \name{expr:\-Entity\-Class}. Entity attribute cardinalities can be only from 0 to 1, even if this attribute is a set of values. In this case the attribute type is an external aggregated data\-type, as in:

@@

Any aggregated type in \ifcowl{} is a subclass of \name{expr:SetClass},
\name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}. The latter has one a nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}. Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}. Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:

@@

\subsubsection{Data Section} In a SPF data section each entity is identified with a line number.
%, an entity data\-type and a list of values. 
All entities of \name{Ifc\-Root}-derived types are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} is a GUID as a hexadecimal string (not including the character \name{\$}). Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format like \name{LINE\_\textless{}lineNumber\textgreater{}} (in a user selected namespace). Note that line numbers are used only for linking entities inside a SPF, and they are unstable. Naming is required when the dataset is used together with OWL 2 EL/QL as these profiles do not allow anonymous individuals. Below is an example of IFC data exported to an RDF dataset compatible with OWL 2 RL:

>>>>

\section{Generating ifcOWL Ontologies from the IFC Schema}
\label{sec:ifcOWL}
This section describes the derivation of the ifcOWL ontologies from the IFC schema. The schema defines series of datatypes, functions and rules by using EXPRESS data definition language (ISO10303-11 \cite{schenck1994information}).

\subsection{Basic Principles}

Layer \ifcsimple{} has been designed to be compatible with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other Semantic Web tools. The other layers \standard{} and \extended{} are compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports many important representational constructs of EXPRESS, such as enumerations of individuals, union classes and anonymous individuals \cite{motik2012owl}. All \ifcowl{} layers include all type definitions and hierarchies so that any \ifcrdf{} datasets converted from IFC data files could reuse types without defining new ones.

The function and rule definitions, type constraints and derived attributes specified in the IFC schema after keywords \name{FUNCTION}, \name{RULE}, and \name{WHERE} are not translated into any \ifcowl{} layer either because of their procedural nature, or the complexity of the resulting constraints.

@@

All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g. \name{IFC4\_ADD1}):

@@

Types \name{STRING} and \name{BINARY} data\-types may be specified with a size parameter: \name{STRING(max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}. In layers \simple{} and \standard{} the size should be ignored, while in layer \extended{} it can be declared:

@@

The enumeration values cannot be defined as literals because enumerations of literals are not supported by OWL 2 RL. Two different approaches are used because enumerations with more than one individual or literal are not supported in OWL 2 EL or QL. The fact that some values -- for instance, \name{ifc:NOT\-DE\-FI\-NED} -- belong to different enumeration data\-types should not cause any problems since any enumeration value is always specified with an exact type.

@@

\subsubsection{Select Datatypes} represent unions of other data\-types. A particular member of a select data\-type can be an entity, or a defined or a select data\-type. For instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}. A select data\-type is converted to a subclass of \name{expr:\-Select\-Class}. Its particular data\-types are defined as subclasses of the select class, but in different ways according to the layer:

@@

Defined data\-types are declared as \name{owl:Class}es using boxing technique because if they were \name{rdfs:Datatype}s they could not be be united with other \name{owl:Class}es for creating select data\-types. In addition, values of defined data\-types must be qualified with type names. For example, a \name{xsd:double} value with type \name{IfcLengthMeasure} is not the same as with type \name{IfcMassMeasure} although these types belong to select class \name{IfcMeasureValue} and defined as equal to \name{REAL}.

In IFC type \name{Ifc\-Time\-Stamp} is defined as \name{INTEGER} which represents so-called Unix time value. However, it was decided that these values must be converted to standard datetime values during reading IFC files. Thus, \name{ifc:Ifc\-Time\-Stamp} wraps a \name{xsd:date\-Time} value instead of \name{xsd:integer}.

@@

\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages. Each type may have one super\-type. An attribute is defined with a data\-type, cardinalities and can be marked as optional. An entity data\-type should be converted to an \name{owl:\-Class} which is subclass of its superclass (if any) or \name{expr:\-Entity\-Class}. It was decided to make all entity properties be functional (the maximum cardinality is 1).

Entity attribute cardinalities can be only from 0 to 1, even if this attribute is a set of values. In this case the attribute type is an external aggregated data\-type, as in:

@@

Any aggregated type in \ifcowl{} is a subclass of \name{expr:SetClass},
\name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}. The latter has one nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}. Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}. Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:

@@

\subsubsection{Data Section} In a SPF data section each entity is identified with a line number.
%, an entity data\-type and a list of values. 
All entities of \name{Ifc\-Root}-derived types have a GUID in IFC. They are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} is in the hexadecimal string representation (not including the character \name{\$}). Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\textless{}lineNumber\textgreater{}} (in a user defined namespace). Note that line numbers are used only for linking entities inside a SPF, and they are unstable. Naming is required when the dataset is used together with OWL 2 EL/QL as these profiles do not allow anonymous individuals. Below is an example of IFC data exported to an RDF dataset compatible with OWL 2 RL:
